Flow-Based Cost Allocation in Linear Optimized Power Systems are non-unique but localizable

### 2020-08-11 paper/flow_based.tex
fix non extendable SUs to extendable

% allocation quantities
\newcommand{\allocatepeer}[1][s, n]{A_{#1,t}}
\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
\newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocatecapexflow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateopex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateemissioncost}[1][n]{\mathcal{E}_{#1,t}}

@@

Now, assume a network with generators $s$, transmission lines $\ell$ and storage units $r$. Each asset $i = \{s, \ell, r\}$ adds an term for OPEX and a term for CAPEX to the total system cost $\totalcost$. The operational price for an asset $i$ is given by $o_i$, the capital price for one unit capacity expansion by $c_i$. For example the OPEX occasioned by generator $s$ is given by 
\begin{align}
    \opexgeneration_s = \sum_t \operationalpricegeneration \, \generation 
    \label{eq:opexgeneration}
\end{align}
where $\generation$ denotes its production at time $t$. It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme} for a full cost allocation of OPEX on production. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
\begin{align}
    \allocateopex[n \rightarrow s] &= 
   \operationalpricegeneration \,  \allocatepeer
\label{eq:allocate_opexGeneration_detailed}
\end{align}
The quantity  
\begin{align}
 \allocatepeer = \pdv{\generation}{\demand} \, \demand
 \label{eq:allocate_peer}
\end{align}
can be considered as a power allocation which indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
It follows naturally that, in the same manner, the scheme can be applied to the OPEX on the flow $\flow$ and storage unit dispatch $\storagedispatch$. As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow. The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation. Note, in all cases, the allocation $\allocatestate$ remains undefined which we will resolve later.

@@

The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
\begin{align}
\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
\label[constraint]{eq:lower_generation_capacity_constraint}
\end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.  As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$, 
\begin{align}
 \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
 \label{eq:no_profit_capex_generation}
\end{align}
This representation connects the CAPEX with the operational state of generator $s$ and matches the needed form in \cref{eq:cost_decomposition} for using the allocation scheme. The resulting the CAPEX allocation is given by
\begin{align}
 \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocatepeer
 \label{eq:allocate_capexGeneration_detailed}
\end{align}
How does this allocation behave? According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not. If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding. Thus it is these times steps which push investments in $\capacitygeneration$. If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary. 
When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.

@@

We us the result of Appendix B.3 in \cite{brown_decreasing_2020} which states that a storage recovers its capital (here we added the operational) costs from aligning dispatch and charging to the LMP, thus 
\begin{align}
    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
\end{align}
where $\incidencestorage$ is one if storage $s$ is at node $n$ and zero otherwise. Using \cref{eq:lagrangian_storage_dispatch}, we can reformulate to 
\begin{align}
    \capitalpricestorage \, \capacitystorage = \sum_t \incidencestorage \left( (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch \right) \storagedispatch 
\end{align}

@@

% Assume the cost-minimal power system includes multiple generators whose dispatch $\generation$ is optimized. The cost term added to $\totalcost$ combines the OPEX for all generators $s$ over the whole time span
% \begin{align}
%     \opexgeneration =  \sum_{s,t} \operationalpricegeneration \, \generation
% \end{align}   
% where $\operationalpricegeneration$ denotes the fixed operational price for generator $s$ to produce 1~MW of electricity. The full cost allocation for the operation of generator $s$, arises from replacing 
% \begin{subequations}
% \label[relation]{eq:opex_allocation}
% \begin{align}
%     i &\Rightarrow s \\
%     \costfactor &\Rightarrow \operationalpricegeneration \\
%     \state &\Rightarrow \generation 
% \end{align}
% \end{subequations}
% % $i$ by $s$, $\costfactor$ by $\operationalpricegeneration$, $\cost$ by $\opexgeneration$ and $\state$ by $\generation$
% in \cref{eq:general_scheme}. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
% \begin{align}
%     \allocateopex[n \rightarrow s] &= 
%    \operationalpricegeneration \,  \allocatepeer
% \label{eq:allocate_opexGeneration_detailed}
% \end{align}
% where the quantity  
% \begin{align}
%  \allocatepeer = \pdv{\generation}{\demand} \, \demand
%  \label{eq:allocate_peer}
% \end{align}
% obtained through \cref{eq:allocate_state}, can be considered as a power allocation. It indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
% From \cref{eq:allocate_state_constraint} it follows that all the power allocations must sum up to the power produced by generator $s$, \ie
% \begin{align}
% \generation = \sum_n \allocatepeer
% \label{eq:allocate_peer_constraint}
% \end{align}
% % 
% % 
% The cost allocation $\allocateopex[n \rightarrow s]$ assumes peer-to-peer tradings between producers and consumers for fixed production prices $\operationalpricegeneration$.
% Hence, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs. In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$, 
% \begin{align}
% \sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
% \label{eq:no_profit_opex}
% \end{align}

@@

Given a fix price for emissions in CO$_2$ equivalents $\emissionprice$, the cost term for emission adds up to 
\begin{align}
 \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
\end{align}
where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
We use the same set of equation as given through \cref{eq:general_scheme,eq:opex_allocation} but update the cost factor to $\costfactor \Rightarrow \emissionprice \, \emission$. For consumers at bus $n$ at time $t$ this leads to the emission cost of 
\begin{align}
 \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocatepeer
\end{align}
associated with generator $s$.

@@

The presented cost allocation suits for any type of topology and network setup. But so far, the question of how $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open. We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:allocate_state_constraint}). \\
% 
In fact, it turns out that ones, the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
In order to show this, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.  Further let $\incidencegenerator$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise. The linear power flow equation can be written as 
\begin{align}
 \flow  = \sum_m \ptdf[m] \left( \nodalgeneration[m] - \demand[m] \right)  
\end{align}
where the nodal generation $\nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation $ combines the production of all generators attached to bus $m$. Taking the derivative with respect to the demand, leads to 
\begin{align}
 \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer[m \rightarrow n]  - \delta_{n,m} \demand \right) 
 \label{eq:allocate_peer_to_allocate_flow}
\end{align}
where $\allocatepeer[m \rightarrow n] = \sum_s \incidencegenerator \, \allocatepeer$ combines all power allocations of the source $m$. \Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocatepeer$ directly determines the flow allocation on line $\ell$.

This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocatepeer$. In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$. Unfortunately, the solution is non-unique and requires further assumptions. Established flow allocation schemes approach this problem from different directions. Principally two options exist \textit{what} is allocated 
% 
\begin{enumerate}
\item gross power injections \label{gross}
\item net power injections \label{net}
\end{enumerate}
% 
Further it is important \textit{what assumptions} define the allocation, \ie what method is used to define the pairs of sources and sinks. The three suitable approaches we present here are
% 
\begin{enumerate}[label=\alph*., ref=\alph*]
\item Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} which assumes
that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010-1} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power decifit or excess of the regions. \label{fbmc}
\end{enumerate}
% 
We show the mathematical formulation for all combinations \ref{ebe}\ref{gross} - \ref{fbmc}\ref{net} in \cref{sec:gross_ebe,sec:net_ebe,sec:gross_ap,sec:net_ap}.
% 
Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $n$ with $\nodalgeneration \le \nodaldemand$ is not assigned to other buses, only to $n$. 
Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocatepeer$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$. Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \generation \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.

@@

\begin{align}
    \allocatepeer[s \rightarrow n] = \dfrac{\generation}{\sum_s \generation} \demand 
\end{align}

@@

Allocating net power injections using the EBE methods leads to the same result as the Marginal Participation (MP) \cite{rudnick_marginal_1995}  algorithm when allocating to consumers only, see \cite{hofmann_flow_2020-1} for further insight. We calculate it by setting 
\begin{align}
\allocatepeer[m \rightarrow n] &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
\label{eq:mp_slack}
\end{align}
where 
\begin{itemize}
%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
\item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
\item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
\item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
\item $\gamma_t = \left( \sum_n \netproduction\right) ^{-1} = \left( \sum_n \netconsumption\right) ^{-1}$ is the inverse of the total injected/extracted power at time $t$.
\end{itemize}

The allocation $\allocatepeer$ from generator $s$ to $n$, is given by multiplying $\allocatepeer[m \rightarrow n]$ with the nodal share $\generation / \nodalgeneration$.

@@

\begin{align}
\InverseAPInjection_t = \left( \diag{\injectionM^+} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( \flow \right)  \incidence$. Then the distributed slack for time step $t$ is given by
\begin{align}
\allocatepeer[m \rightarrow n] = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
\end{align}

@@

We use the same allocation as in \cref{sec:net_ap} but replace the net nodal production $\netproduction$ by the gross nodal production $\nodalgeneration$ which leads to  
\begin{align}
\InverseAPInjection_t = \left( \diag{g} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
The distributed slack is for time step $t$ is then given by
\begin{align}
\allocatepeer[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
\end{align}

>>>>

% allocation quantities
\newcommand{\allocategeneration}[1][s, n]{A_{#1,t}}
\newcommand{\allocatepeer}[1][m \rightarrow n]{A_{#1,t}}
\newcommand{\allocateflow}[1][n]{A_{\ell,#1,t}}
\newcommand{\allocatetransaction}[1][s \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocatestate}[1][i, n]{A_{#1,t}}
\newcommand{\allocatecapexgeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocatecapexflow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateopex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateemissioncost}[1][n]{\mathcal{E}_{#1,t}}

@@

Now, assume a network with generators $s$, transmission lines $\ell$ and storage units $r$. Each asset $i = \{s, \ell, r\}$ adds an term for OPEX and a term for CAPEX to the total system cost $\totalcost$. The operational price for an asset $i$ is given by $o_i$, the capital price for one unit capacity expansion by $c_i$. For example the OPEX occasioned by generator $s$ is given by 
\begin{align}
    \opexgeneration_s = \sum_t \operationalpricegeneration \, \generation 
    \label{eq:opexgeneration}
\end{align}
where $\generation$ denotes its production at time $t$. It becomes clear that \cref{eq:opexgeneration} matches the form of \cref{eq:cost_decomposition} which allows us to use the above presented scheme in \cref{eq:general_scheme} for a full cost allocation of OPEX on production. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
\begin{align}
    \allocateopex[n \rightarrow s] &= 
   \operationalpricegeneration \,  \allocategeneration
\label{eq:allocate_opexGeneration_detailed}
\end{align}
The quantity  
\begin{align}
 \allocategeneration = \pdv{\generation}{\demand} \, \demand
 \label{eq:allocate_peer}
\end{align}
can be considered as a power allocation which indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
It follows naturally that, in the same manner, the scheme can be applied to the OPEX on the flow $\flow$ and storage unit dispatch $\storagedispatch$. As we assume a bidirectional flow on line $\ell$, the OPEX is set proportional to the absolute value of the flow. The upper section in \cref{tab:cost_allocation_map} shows the mapping of variables to \cref{eq:general_scheme} in order to define the full OPEX allocation. Note, in all cases, the allocation $\allocatestate$ remains undefined which we will resolve later.

@@

The nominal capacity $\capacitygeneration$ constrains the generation $\generation$ in the form of 
\begin{align}
\generation - \generationpotential \capacitygeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{s,t} 
\label[constraint]{eq:upper_generation_capacity_constraint}\\ 
- \generation &\le 0 \resultsin{\mulowergeneration} \Forall{s,t} 
\label[constraint]{eq:lower_generation_capacity_constraint}
\end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. At a cost-optimum, these two constraints yield the shadow prices $\muuppergeneration$ and $\mulowergeneration$.  As shown in \cite{brown_decreasing_2020} and in detailed in \cref{sec:zero_profit_generation}, over the whole time span, the CAPEX for generator $s$ is payed back by the production $\generation$ times the shadow price $\muuppergeneration$, 
\begin{align}
 \capexgeneration_s = \capitalpricegeneration \capacitygeneration = \sum_t \muuppergeneration \,  \generation 
 \label{eq:no_profit_capex_generation}
\end{align}
This representation connects the CAPEX with the operational state of generator $s$ and matches the needed form in \cref{eq:cost_decomposition} for using the allocation scheme. The resulting the CAPEX allocation is given by
\begin{align}
 \allocatecapexgeneration[n \rightarrow s] = \muuppergeneration \, \allocategeneration
 \label{eq:allocate_capexGeneration_detailed}
\end{align}
How does this allocation behave? According to the polluter pays principle, it differentiates between consumers who are `responsible` for investments and those who are not. If $\muuppergeneration$ (in literature often denoted as the Quality of Supply) is bigger than zero, the upper Capacity \cref{eq:upper_generation_capacity_constraint} is binding. Thus it is these times steps which push investments in $\capacitygeneration$. If $\muuppergeneration = 0$, the generation $\generation$ is not bound and investments are not necessary. 
When summing over all CAPEX payments to generator $s$ in \cref{eq:allocate_capexGeneration_detailed} and using \cref{eq:allocate_state_constraint,eq:no_profit_capex_generation}, we see that each generator retrieves exactly the cost that were spent to build the capacity $\capacitygeneration$.

@@

We use the result of Appendix B.3 in \cite{brown_decreasing_2020} which shows that a storage recovers its capital (and operational) costs from aligning dispatch and charging to the LMP, thus 
\begin{align}
    \sum_t \operationalpricestorage \, \storagedispatch + \capitalpricestorage \, \capacitystorage = \sum_t \lmp \incidencestorage \left(\storagedispatch - \storagecharge \right) 
\end{align}
where $\incidencestorage$ is one if storage $s$ is at node $n$ and zero otherwise. Using \cref{eq:lagrangian_storage_dispatch}, we can reformulate to 
\begin{align}
    \capitalpricestorage \, \capacitystorage = \sum_t \incidencestorage \left( (\efficiencydispatch )^{-1} \mustateofcharge - \mulowerstoragedispatch \right) \storagedispatch 
\end{align}

@@

% Assume the cost-minimal power system includes multiple generators whose dispatch $\generation$ is optimized. The cost term added to $\totalcost$ combines the OPEX for all generators $s$ over the whole time span
% \begin{align}
%     \opexgeneration =  \sum_{s,t} \operationalpricegeneration \, \generation
% \end{align}   
% where $\operationalpricegeneration$ denotes the fixed operational price for generator $s$ to produce 1~MW of electricity. The full cost allocation for the operation of generator $s$, arises from replacing 
% \begin{subequations}
% \label[relation]{eq:opex_allocation}
% \begin{align}
%     i &\Rightarrow s \\
%     \costfactor &\Rightarrow \operationalpricegeneration \\
%     \state &\Rightarrow \generation 
% \end{align}
% \end{subequations}
% % $i$ by $s$, $\costfactor$ by $\operationalpricegeneration$, $\cost$ by $\opexgeneration$ and $\state$ by $\generation$
% in \cref{eq:general_scheme}. As a result we obtain the contribution of $\demand$ to the OPEX at generator $s$, given by 
% \begin{align}
%     \allocateopex[n \rightarrow s] &= 
%    \operationalpricegeneration \,  \allocategeneration
% \label{eq:allocate_opexGeneration_detailed}
% \end{align}
% where the quantity  
% \begin{align}
%  \allocategeneration = \pdv{\generation}{\demand} \, \demand
%  \label{eq:allocate_peer}
% \end{align}
% obtained through \cref{eq:allocate_state}, can be considered as a power allocation. It indicates the power that is produced by generator $s$ and consumed at node $n$ at time $t$.
% From \cref{eq:allocate_state_constraint} it follows that all the power allocations must sum up to the power produced by generator $s$, \ie
% \begin{align}
% \generation = \sum_n \allocategeneration
% \label{eq:allocate_peer_constraint}
% \end{align}
% % 
% % 
% The cost allocation $\allocateopex[n \rightarrow s]$ assumes peer-to-peer tradings between producers and consumers for fixed production prices $\operationalpricegeneration$.
% Hence, consumers at bus $n$ which retrieve power from a generators $s$, directly  compensate its operational costs. In this way,  generator $s$ retrieves the exact amount of money from consumers that it spends on its operation at time $t$, 
% \begin{align}
% \sum_{n} \allocateopex[n \rightarrow s] = \operationalpricegeneration \, \generation
% \label{eq:no_profit_opex}
% \end{align}

@@

Given a fix price for emissions in CO$_2$ equivalents $\emissionprice$, the cost term for emission adds up to 
\begin{align}
 \emissioncost = \emissionprice \, \sum_s  \emission \, \generation
\end{align}
where $\emission$ denotes the emission factor in tonne-CO$_2$ per \megawatthour for generator $s$.
We use the same set of equation as given through \cref{eq:general_scheme,eq:opex_allocation} but update the cost factor to $\costfactor \Rightarrow \emissionprice \, \emission$. For consumers at bus $n$ at time $t$ this leads to the emission cost of 
\begin{align}
 \allocateemissioncost[n \rightarrow s] = \emissionprice \, \emission \, \allocategeneration
\end{align}
associated with generator $s$.

@@

The presented cost allocation suits for any type of topology and network setup. But so far, the question of how $\allocatestate$ for generators $s$, lines $\ell$ and storages $r$ are defined was left open. We recap that all rely on derivatives of $\generation$, $\flow$ and $\storagedispatch$  with respect to the nodal demand $\demand$ (see \cref{eq:allocate_state_constraint}). \\
% 
In fact, it turns out that ones, the peer-to-peer allocations $\allocatepeer$ are defined, the power flow allocations $\allocateflow$ are determined by the linear power flow equations.
In order to show this, let $\ptdf$ denote Power Transfer Distribution Factors (PTDF) giving the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$.  Further let $\incidencegenerator$ be 1 if generator $s$ is attached to bus $n$ and zero otherwise. The linear power flow equation can be written as 
\begin{align}
 \flow  = \sum_m \ptdf[m] \left( \nodalgeneration[m] - \demand[m] \right)  
\end{align}
where the nodal generation $\nodalgeneration[m] = \sum_s \incidencegenerator[m] \, \generation $ combines the production of all generators attached to bus $m$. Taking the derivative with respect to the demand, leads to 
\begin{align}
 \allocateflow = \pdv{\flow}{\demand} \demand = \sum_m \ptdf[m] \left( \allocatepeer  - \delta_{n,m} \demand \right) 
 \label{eq:allocate_peer_to_allocate_flow}
\end{align}
where $\allocatepeer = \sum_s \incidencegenerator \, \allocategeneration$ combines all power allocations of the source $m$. \Cref{eq:allocate_peer_to_allocate_flow} shows that the peer-to-peer allocation $\allocategeneration$ directly determines the flow allocation on line $\ell$.

This leaves us with only solving \cref{eq:allocate_peer_constraint} for $\allocategeneration$. In fact, we have to answer the question of how much power supply does generator $s$ deliver to bus $n$. Unfortunately, the solution is non-unique and requires further assumptions. Established flow allocation schemes approach this problem from different directions. Principally two options exist \textit{what} is allocated 
% 
\begin{enumerate}
\item gross power injections \label{gross}
\item net power injections \label{net}
\end{enumerate}
% 
Further it is important \textit{what assumptions} define the allocation, \ie what method is used to define the pairs of sources and sinks. The three suitable approaches we present here are
% 
\begin{enumerate}[label=\alph*., ref=\alph*]
\item Equivalent Bilateral Exchanges (EBE) \cite{galiana_transmission_2003} which assumes
that every producer supplies every consumer proportional to its share in the total consumption. \label{ebe} 
\item Average Participation (AP) \cite{bialek_tracing_1996,achayuthakan_electricity_2010-1} which traces the flow from producer to consumer following the law of proportional sharing. \label{ap}
\item Flow Based Market Coupling (FBMC) which uses zonal PTDF for allocating power within predefined regions. The interregional exchange is only allocating net power decifit or excess of the regions. \label{fbmc}
\end{enumerate}
% 
We show the mathematical formulation for all combinations \ref{ebe}\ref{gross} - \ref{fbmc}\ref{net} in \cref{sec:gross_ebe,sec:net_ebe,sec:gross_ap,sec:net_ap}.
% 
Principally, type \ref{net} leads to less P2P trades then type \ref{gross} as power from a bus $n$ with $\nodalgeneration \le \nodaldemand$ is not assigned to other buses, only to $n$. 
Further, as literature has often pointed out, the EBE principal \ref{ebe} does not suit for large networks where remote buses would interconnect in the same way as buses in close vicinity \cite{gil_multiarea_2005}. The AP based type \ref{ap} tackles this problem by restricting P2P trades to those which are traceable when applying the proportional sharing principal. Therefore $\allocategeneration$ denotes that part of power produced by bus $m$ which, when only following in the direction of $\flow$, ends up at bus $n$. Type \ref{fbmc} further allows to control the regions or market zones which are netted out in a first calculation. If in a region $R$ the generation undercuts the demand, $\sum_{n \in R} \generation \le \sum_{n \in R} \demand$, none of the inner-regional generation is assigned to other regions.

@@

\begin{align}
    \allocategeneration = \dfrac{\generation}{\sum_s \generation} \demand 
\end{align}

@@

Allocating net power injections using the EBE methods leads to the same result as the Marginal Participation (MP) \cite{rudnick_marginal_1995}  algorithm when allocating to consumers only, see \cite{hofmann_flow_2020-1} for further insight. We calculate it by setting 
\begin{align}
\allocatepeer &=  \delta_{m,n}\,\selfconsumption[m] + \gamma_t \, \netconsumption  \, \netproduction[m]
\label{eq:mp_slack}
\end{align}
where 
\begin{itemize}
%  \item $\injection = \left( \nodalgeneration - \nodaldemand \right) $ denotes the nodal injection
\item $\netproduction = \text{min}\left( \nodalgeneration - \nodaldemand , 0 \right) $ denotes the nodal net production 
\item $\netconsumption = \text{min}\left( \nodaldemand  - \nodalgeneration, 0 \right)$ denotes the nodal net consumption
\item $\selfconsumption = \text{min}\left( \netproduction, \netconsumption \right)$ the denotes  nodal self-consumption. That is the power generated and at the same time consumed at node $n$ and 
\item $\gamma_t = \left( \sum_n \netproduction\right) ^{-1} = \left( \sum_n \netconsumption\right) ^{-1}$ is the inverse of the total injected/extracted power at time $t$.
\end{itemize}

The allocation $\allocategeneration$ from generator $s$ to $n$, is given by multiplying $\allocatepeer$ with the nodal share $\generation / \nodalgeneration$.

@@

\begin{align}
\InverseAPInjection_t = \left( \diag{\injectionM^+} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
where $\DirectedIncidence^-$ is the negative part of the directed Incidence matrix $\DirectedIncidence_{n,\ell} = \text{sign}\left( \flow \right)  \incidence$. Then the distributed slack for time step $t$ is given by
\begin{align}
\allocatepeer = \InverseAPInjection_{m,n,t} \, \netproduction[m] \, \netconsumption
\end{align}

@@

We use the same allocation as in \cref{sec:net_ap} but replace the net nodal production $\netproduction$ by the gross nodal production $\nodalgeneration$ which leads to  
\begin{align}
\InverseAPInjection_t = \left( \diag{g} + \DirectedIncidence^- \diag{\flowM} \, \incidenceM \right)_t^{-1} 
\end{align}
The distributed slack is for time step $t$ is then given by
\begin{align}
\allocategeneration[s \rightarrow m] = \InverseAPInjection_{m,n} \, \generation \, \demand
\end{align}
